Internal models of type theory

### 2013-10-31 article-models-defn.tex
back to \jterm for terms

\begin{rmk}
Every term $\jtermt{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
\begin{equation*}
\inference{\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtype{\ctxext{\Gamma}{A}}{f^\ast Q}}
\qquad
\inference{\jtermt{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast g}}
\end{equation*}
\end{rmk}

\begin{defn}
A term $f:\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$ is said to be a trivial
cofibration if we can infer
\begin{equation*}
\inference{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermt{\ctxext{\Gamma}{B}}{Q}{\tilde{t}}}\qquad
\inference{\jtermt{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermteq{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast \tilde{t}}{t}}
\end{equation*}
{\color{red}This statement should be reformulated so that it only involves a single judgment...
but I don't see directly how to do that.}
\end{defn}

@@

\begin{infarray}{c}
\inference{\jtermt{\Gamma}{A}{x}}{\jtype{\ctxext{\Gamma}{A}}{\eqtype{x}}}\\
\inference{\jtermt{\Gamma}{A}{x}}{\jtermt{\Gamma}{\subst{x}{\eqtype{x}}}{\refl{x}}}\\
\inference{\jtype{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}\qquad\jtermt{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
          {\jtermt{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}{\ind{\eqtype{x}}(d)}}\\
\inference{\jtype{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}\qquad\jtermt{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
          {\jtermteq{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{\subst{\refl{x}}{{x}{P}}}{\subst{\refl{x}}{{x}{\ind{\eqtype{x}}(d)}}}{d}}
\end{infarray}

We get something that looks like Paulin-Mohring equality. But we should be able to use it to
show that every $\jtermt{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.

For the following conjecture, note that if we have $\jtermt{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$
Then we may consider the type $\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}
{\ctxwk{\Gamma}{\idtypevar{\Delta}}}$ and we may substitute $f$ to obtain the type
\begin{equation*}
\jtype{\ctxext{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}}
{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}
\end{equation*}
Note that we have the judgmental equality
\begin{equation*}
\jtypeeq{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}{\ctxwk{\Gamma}{\Delta}}
\end{equation*}
so we obtain the type
\begin{equation*}
\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}
\end{equation*}
We also have $\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}$, so we
may ask for a term of type $\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}$
in context $\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}$ and we can ask ourselves the question
whether this is a trivial cofibration. This is the content of the following conjecture.

\begin{conj}
There is a term
\begin{equation*}
\jtermt{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}{\eqtoid{f}}
\end{equation*}
Moreover, $\eqtoid{f}$ is a trivial cofibration.
\end{conj}

\begin{proof}
We use the induction principle of $\eqtype{f}$ to construct $\eqtoid{f}$. Thus, we
have to find a term
\begin{equation*}
\jtermt
  {\Gamma}
  {\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}}}
  {\subst{\refl{f}}{{f}{\eqtoid{f}}}}
\end{equation*}
Note that we have the judgmental equalities
\begin{align*}
\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}}
& \jdeq \subst{f}{{\ctxwk{{\Gamma}{\Delta}}{\refl{f}}}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}} \\
& \jdeq \subst{f}{\ctxwk{\Gamma}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}
\end{align*}
where we find the term $\subst{f}{\ctxwk{\Gamma}{\subst{f}{\ctxwk{\Gamma}{\refl{\Delta}}}}}$.
\end{proof}

\begin{conj}
Suppose we have a type $\eqtype{x}$ for every term $\jtermt{\Gamma}{A}{x}$. Then
every term $\jtermt{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.
\end{conj}

\begin{proof}
Suppose that $Q$ is a type in context $\Delta$. We want to show that
\begin{equation*}
\inference{\jtermt{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{g}}{\jtermt{\Delta}{Q}{\tilde{g}}}
\end{equation*}
Let $g$ be a term of $\subst{f}{\ctxwk{\Gamma}{Q}}$ in context $\Gamma$. We get
the term
\begin{equation*}
\jtermt{\Gamma}{\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{{\Gamma}{Q}}}}}{\subst{\refl{f}}{\ctxwk{\eqtype{f}}{g}}}
\end{equation*}
which is judgmentally equal to $g$; the types

For any term $\jtermt{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{g}$ we get a term 
$\jtermt{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}$.

$\jtermt{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\Gamma}{Q}}{\ind{\eqtype{f}}(d)}$.
\end{proof}

@@

\begin{infarray}{c}
\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{\ctxext{\Gamma}{A}}{\Gamma}}}{\eqtype{\proj1^A}}}\\
\inference{\jtype{\Gamma}{A}}{\jtermt{\ctxext{\Gamma}{A}}{\subst{\proj1^A}{\eqtype{\proj1^A}}}{\refl{\proj1^A}}}
\end{infarray}

>>>>

\begin{rmk}
Every term $\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
\begin{equation*}
\inference{\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtype{\ctxext{\Gamma}{A}}{f^\ast Q}}
\qquad
\inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast g}}
\end{equation*}
\end{rmk}

\begin{defn}
A term $f:\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$ is said to be a trivial
cofibration if we can infer
\begin{equation*}
\inference{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jterm{\ctxext{\Gamma}{B}}{Q}{\tilde{t}}}\qquad
\inference{\jterm{\ctxext{\Gamma}{A}}{f^\ast Q}{t}}{\jtermeq{\ctxext{\Gamma}{A}}{f^\ast Q}{f^\ast \tilde{t}}{t}}
\end{equation*}
{\color{red}This statement should be reformulated so that it only involves a single judgment...
but I don't see directly how to do that.}
\end{defn}

@@

\begin{infarray}{c}
\inference{\jterm{\Gamma}{A}{x}}{\jtype{\ctxext{\Gamma}{A}}{\eqtype{x}}}\\
\inference{\jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{\subst{x}{\eqtype{x}}}{\refl{x}}}\\
\inference{\jtype{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}\qquad\jterm{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
          {\jterm{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}{\ind{\eqtype{x}}(d)}}\\
\inference{\jtype{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{P}\qquad\jterm{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}}
          {\jtermeq{\ctxext{{\Gamma}{A}}{\eqtype{x}}}{\subst{\refl{x}}{{x}{P}}}{\subst{\refl{x}}{{x}{\ind{\eqtype{x}}(d)}}}{d}}
\end{infarray}

We get something that looks like Paulin-Mohring equality. But we should be able to use it to
show that every $\jterm{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.

For the following conjecture, note that if we have $\jterm{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$
Then we may consider the type $\jtype{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}
{\ctxwk{\Gamma}{\idtypevar{\Delta}}}$ and we may substitute $f$ to obtain the type
\begin{equation*}
\jtype{\ctxext{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}}
{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}
\end{equation*}
Note that we have the judgmental equality
\begin{equation*}
\jtypeeq{\Gamma}{\subst{f}{\ctxwk{\Gamma}{{\Delta}{\Delta}}}}{\ctxwk{\Gamma}{\Delta}}
\end{equation*}
so we obtain the type
\begin{equation*}
\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}
\end{equation*}
We also have $\jtype{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}$, so we
may ask for a term of type $\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}$
in context $\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}$ and we can ask ourselves the question
whether this is a trivial cofibration. This is the content of the following conjecture.

\begin{conj}
There is a term
\begin{equation*}
\jterm{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}{\eqtoid{f}}
\end{equation*}
Moreover, $\eqtoid{f}$ is a trivial cofibration.
\end{conj}

\begin{proof}
We use the induction principle of $\eqtype{f}$ to construct $\eqtoid{f}$. Thus, we
have to find a term
\begin{equation*}
\jterm
  {\Gamma}
  {\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}}}
  {\subst{\refl{f}}{{f}{\eqtoid{f}}}}
\end{equation*}
Note that we have the judgmental equalities
\begin{align*}
\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}}
& \jdeq \subst{f}{{\ctxwk{{\Gamma}{\Delta}}{\refl{f}}}{\ctxwk{\eqtype{f}}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}} \\
& \jdeq \subst{f}{\ctxwk{\Gamma}{\subst{f}{\ctxwk{\Gamma}{\idtypevar{\Delta}}}}}
\end{align*}
where we find the term $\subst{f}{\ctxwk{\Gamma}{\subst{f}{\ctxwk{\Gamma}{\refl{\Delta}}}}}$.
\end{proof}

\begin{conj}
Suppose we have a type $\eqtype{x}$ for every term $\jterm{\Gamma}{A}{x}$. Then
every term $\jterm{\Gamma}{\ctxwk{\Gamma}{\Delta}}{f}$ is a trivial cofibration.
\end{conj}

\begin{proof}
Suppose that $Q$ is a type in context $\Delta$. We want to show that
\begin{equation*}
\inference{\jterm{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{g}}{\jterm{\Delta}{Q}{\tilde{g}}}
\end{equation*}
Let $g$ be a term of $\subst{f}{\ctxwk{\Gamma}{Q}}$ in context $\Gamma$. We get
the term
\begin{equation*}
\jterm{\Gamma}{\subst{\refl{f}}{{f}{\ctxwk{\eqtype{f}}{{\Gamma}{Q}}}}}{\subst{\refl{f}}{\ctxwk{\eqtype{f}}{g}}}
\end{equation*}
which is judgmentally equal to $g$; the types

For any term $\jterm{\Gamma}{\subst{f}{\ctxwk{\Gamma}{Q}}}{g}$ we get a term 
$\jterm{\Gamma}{\subst{\refl{x}}{{x}{P}}}{d}$.

$\jterm{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\eqtype{f}}}{\ctxwk{\Gamma}{Q}}{\ind{\eqtype{f}}(d)}$.
\end{proof}

@@

\begin{infarray}{c}
\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{\ctxext{\Gamma}{A}}{\Gamma}}}{\eqtype{\proj1^A}}}\\
\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\subst{\proj1^A}{\eqtype{\proj1^A}}}{\refl{\proj1^A}}}
\end{infarray}
